### 2023-01-11 paper/writeup.tex
added discrepancy function

\begin{algorithm}[!ht]

\input{$(x,ts_1,ts_2)$}
\output{discrepancy number (some float value)}

$sim \gets np.zeros(ts2.num_nodes)$\;

\caption{Node Tree Discrepancy}\label{ndisc}
\end{algorithm}

>>>>

\par Now with an algorithm in place
 we need a metric to discern
 if trees modified with edge extend
 in fact correctly infer haplotypes in our original tree sequence.
 We have thus constructed a function which counts the differences 
 between trees on a pair of tree sequences. 
 To count all the discrepancies we must begin by counting differences for each node in the tree sequence.

 Let $TS^1=(\{T_i^1\},N^1,E^1,L^1)$ and $TS^2=(\{T^2_j\},N^2,E^2,L^2)$ be tree sequences.
 Suppose that $x\in N^1$ and $y\in N^2$ are nodes in each tree sequence.
 We denote the \textit{set of descendant samples} of $x$ at position $i$ in the genome as $D_{TS^1}(x,i)$.
 Now for a pair of nodes $(x,y)\in TS^1\times TS^2$ 
 we can measure how much they agree along the genome.
 We do so with the following function
 $$\rho_{(TS^1,TS^2)}(x,y)=\frac{1}{L}\sum_{i=0}^{L-1}[D_{TS^1}(x,i)=D_{TS^2}(y,i)],$$
 where $L=\min\{L^1,L^2\}$ and $[\cdot]$ is the characteristic function. 
 We see $\rho$ measures the fraction of the tree sequence
 along which $x$ and $y$ have exactly the same descendant set. 
 For each node $x\in N^1$ we can use $\rho$ to find its global best matching node $y\in N^2$. 
 This node $y$ should not just be a maximum of $\rho(x,\cdot)$ but it should also have a branch length similar to $x$. 
 This means our node $y$ should minimize both $|t_x-t_y|$ and $1-\rho_{(TS^1,TS^2)}$.
 We then use this $y$ to find count the length of genome which $x$ differs from its best matching node in $TS_2$ with the function
 $$\hat{d}(x,TS^1,TS^2)=\min_y\{|t_x-t_y|(1-\rho_{(TS^1,TS^2)}(x,y))\}.$$
 Here $t_x$ and $t_y$ denote the time of occurrences (ie. branch length) for nodes $x$ and $y$. 
 The function $\hat{d}$ measures the proportion of the genome 
 over which the descendant set of $x$ differs from its best matching node $y$.
 We can now extend this from any node $x\in N^11$ to all of $TS_1$ with the \textbf{discrepancy function}
 $$d(TS_1,TS_2)=\sum_{x\in N_1}\hat{d}(x,TS_1,TS_2).$$
 This function, as we will see in the algorithm below,
 is $|N^1||\{T^1\}|$ as for each tree in $TS_1$ and each node $x\in N^1$ we only need to look at $O(1)$ nodes rather than all nodes in $N^2$.

\begin{algorithm}[!ht]
\SetStartEndCondition{ }{}{}%
\SetKwProg{Fn}{def}{\string:}{}
\SetKwFunction{Range}{range}%%
\SetKw{KwTo}{in}
\SetKwFor{For}{for}{\string:}{}%
\SetKwIF{If}{ElseIf}{Else}{if}{:}{elif}{else:}{}%
\SetKwFor{While}{while}{:}{fintq}%
\SetKw{Break}{break}
\newcommand\forcond{$i$ \KwTo\Range{$n$}}
\SetKwComment{Comment}{}{} 
\DontPrintSemicolon
\LinesNotNumbered
\AlgoDontDisplayBlockMarkers\SetAlgoNoEnd\SetAlgoNoLine%

\input{$(TS_1,TS_2)$}
\output{discrepancy number (some float value)}

\Fn{Node Discrepancy($x,TS_1,TS_2$)}{
$s=0\in \mathbb{Z}^{|N^2|}$ \Comment*[f]{Similarity vector with each entry corresponds to a node in $TS_2$}

\For{$I$, $T^1_k$, $T^2_j$ \KwTo $(L,TS_1,TS_2)$}{
\Comment{Here $I=[a^1_k,a^1_{k+1})\cap [a^2_j,a^2_{j+1})$ is the interval which the two trees $T^1_k$, $T^2_j$ both span over. Here we find the best matching node $y$.}
$D_{T^1_k}(x)\gets D_{TS^1}(x,i)$ where $i\in I$\;
\If{$|D_{T^1_k}(x)| > 0$}{
\If{$|D_{T^1_k}(x)|=1$}{
$y \gets D_{T^1_k}(x)$\;
$m \gets D_{T^1_k}(x)$\;}
\Else{$m \gets $ most recent common anscestor of $D_{T^1_k}(x)$ in $T^1_k$\;
$y\gets $ most recent common anscestor of $D_{T^1_k}(x)$ in $T^2_j$\;}}
\If{$D_{T^2_j}(y)=D_{T^1_k}(x)$}{
$s_y\gets s_y$\;}}
\KwRet{node discrepancy}\;}
\Fn{Discrepancy($TS_1,TS_2$)}{
discrepancy = 0\;
\For{$x\in N^1$}{discrepancy = $\sum_{x\in N^1}{$Node Discrepancy($x,TS_1,TS_2$)}}
\BlankLine
\KwRet{discrepancy}}
\caption{Discrepancy Function}\label{ndisc}
\end{algorithm}
